---
title: "Session 2"
author: "Antoine Mayerowitz"
date: "17/09/2018"
output:
  pdf_document: default
  html_document: default
---

```{r, include: FALSE, warning = FALSE, echo = FALSE, error = TRUE, include=FALSE}
# A cool package manager :
if (!require("pacman")) install.packages("pacman")

# Install and/or load needed packages
pacman::p_load(tidyverse, # Many tools for data handling. Check https://www.tidyverse.org/
               devtools,  # Needed to download ScPoEconometrics package
               rvcheck    # Used to check ScPoEconometrics package version
              )

# Install and/or update and/or load the ScPo package
if (!check_github("ScPoEcon/ScPoEconometrics")$up_to_date) {
   install_github("ScPoEcon/ScPoEconometrics") 
}
library("ScPoEconometrics")
```

# Chapter 1 (continued)
## Matrices
### Definition
```{r}
X = matrix(nrow= 2, ncol=5)
X = matrix(0, nrow= 2, ncol=5)
X = matrix(1:10, nrow= 2, ncol=5)
print(X) # Let's look at X (! Note that we just redefined X three times.)
```
### Subsetting
```{r}
X[1, ] # The first row
X[ ,1] # First column
X[2,3] # 2nd row, 3rd column
```
### Operations
```{r}
X = matrix(1:4, 2, 2)
Y = matrix(4:1, 2, 2)
X * Y # Element wise multiplication
X %*% Y # Matrix multiplication
```
### Task 3
- Create a vector containing 1,2,3,4,5 called v.
- Create the (2,5) matrix m with data (1:10)
- Perform matrix multiplication of m with v. Use the command %*%. What dimension does the output have?
- Why does the command v %*% m not work?
```{r}
v = 1:5
print(v)
m = matrix(1:10, nrow = 2, ncol = 5)
print(m)
dim(m %*% v)
```
## Lists
### Definition
```{r}
myList = list(
  name = "John",
  age= 32,
  matrix = matrix(1:4, 2, 2)
)
```
### Subsetting
```{r}
myList$name
myList[1]
myList[[1]]
```
### Task 4
- Copy and paste the above code for ex_list into your R session. Remember that list can hold any kind of R object. Like…another list! So, create a new list new_list that has two fields: a first field called “this” with string content "is awesome", and a second field called “ex_list” that contains ex_list.
- Accessing members is like in a plain list, just with several layers now. Get the element c from `ex_list` in `new_list`!
- Compose a new string out of the first element in new_list, the element under label this. Use the function paste to print the string 'R is awesome' to your screen.
```{r}
ex_list = list(
  a = c(1, 2, 3, 4),
  b = TRUE,
  c = "Hello!",
  d = function(arg = 42) {print("Hello World!")},
  e = diag(5)
)

new_list = list(
  this = "is awesome",
  ex_list = ex_list
)

new_list$ex_list$c

paste("R", new_list$this)
```

## DataFrame
### Definition
```{r}
data = data.frame(
  x = 1:10,
  y = sample(c("a", "b", "c"), size= 10 , replace= TRUE)
)
```
### Some methods
```{r}
nrow(data) # Number of rows
ncol(data) # Number of columns
names(data) # Variables names
head(data) # Prints a few rows
```
### Example (mtcars)
```{r}
head(mtcars, n=3) # Shows the first 3 rows
mtcars[,"mpg"] # Subset only the `mpg` column
# Subset
mtcars[mtcars$mpg > 20, c("mpg", "cyl")] # Subset lines where mpg > 20 and keep only the columns 'mpg' and 'cyl'
subset(mtcars, subset = mpg > 20, select = c("mpg", "cyl")) # Another way to write it

```
### Task 5
- How many observations are there in mtcars?
- How many variables?
- What is the average value of mpg?
- What is the average value of mpg for cars with more than 4 cylinders, i.e. with cyl>4?
```{r}
nrow(mtcars)
ncol(mtcars)
mean(mtcars$mpg)
mean(subset(mtcars, subset= cyl > 4)$mpg)

# Create new var
data = mtcars[, c("mpg", "cyl" )]
data$myCol = data$mpg / data$cyl
head(data)
```


## Basic Programming
### Variables
```{r}
# TYPE OF DATA :
1 # integer
1.02 # numeric
"Hello" # String
"a" # character
1:2 # vector
  # and also matrix, list, dataframe

# We bind those objects to a variable, which is juste a name, a placeholder, a reference, you name it.
```
### Control Flow
```{r}
# How to treat objects differently based on some characteristics.
# Below we write a if/else statement that turns x into |x|
x = 2
if (x < 0) {
  x = -x
} else {
  x = x 
}
```
### Loops
```{r}
# Iterate along a sequence (can be any sequence)
for (i in c("mangos", "bananas", "apples")){
  print(paste("I love",i))
}
```
### Functions
### Task 6
- Write a for loop that counts down from 10 to 1, printing the value of the iterator to the screen.
- Modify that loop to write “i iterations to go” where i is the iterator
- Modify that loop so that each iteration takes roughly one second. You can achieve that by adding the command Sys.sleep(1) below the line that prints “i iterations to go”.
- Finally, let's create a function called ticking_bomb. it takes no arguments, it's body is the loop you wrote in the preceding question. The only think you should add to the body is a line after the loop finishes, printing “BOOOOM!” with print("BOOOOM!"). You can repeatedly redefine the function in the console, and try it out with ticking_bomb().
```{r}
for (i in 1:10){
  print(i)
}

for (i in 10:1){
  print(paste(i, "iterations to go"))
  Sys.sleep(1)
}

ticking_bomb = function(){
  for (i in 10:1){
    print(paste(i, "iterations to go"))
    Sys.sleep(0.3)
  }
  print('BOOOOM')
}

ticking_bomb() # Don't forget to call the function, and don't forget the parentheses !!
```



